Manuscript progress report

### 2024-02-06 Docs/manuscript.Rmd
figures and tables

---
title: "Manuscript progress report"
author: "Ying Jin"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float: yes
    font: 12pt
    keep_tex: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
    keep_tex: yes
---

@@

bind_cols(fgfpca1, gfosr_l5_1, gfosr_l1_1, adglmm1, .name_repair = "minimal") %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" " = 1, "fGFPCA" = 4, "GFOSR (L=5)" = 4, 
                     "GFOSR (L=1)" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" " = 1, "Maximum observation time" = 16)) %>%
  group_rows(index = c("ISE"= 4, "AUC" = 4))

@@

bind_cols(fgfpca2, gfosr_l5_2, gfosr_l1_2, adglmm2, .name_repair = "minimal") %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" " = 1, "fGFPCA" = 4, "GFOSR (L=5)" = 4, 
                     "GFOSR (L=1)" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" " = 1, "Maximum observation time" = 16)) %>%
  group_rows(index = c("ISE"= 4, "AUC" = 4))

@@

We take 500 subjects for training, 500 for out-of-sample prediction. When using the full sample The debiase step (mgcv::bam) in fGFPCA took a really a long time, and GLMM adaptive only works for intercept-only model. In the debias step, re-evaluated eigenvalues also lost their decreasing nature.

In addition to reduce sample size, maybe we could just drop the debias step and only interpolate the eigenfunctions?

```{r load_nhanes, eval=FALSE}
df_nhanes <- read_rds(here("Data/nhanes_bi.rds"))
load(here("Data/ApplOutput.RData"))
```

@@

```{r auc_fGFPCA_nhanes, eval=FALSE}
window3 <- c(0, 480, 960, 1440)

auc_mat_nhanes <- nhanes_pred_fgfpca %>%
  mutate(window = cut(t, breaks = window3, include.lowest = T)) %>% 
  select(Y, starts_with("pred"), window) %>%
  group_by(window) %>%
  summarise(auc1 = get_auc(Y, pred480),
            auc2 = get_auc(Y, pred960)) %>%
  filter(window != "[0,480]") %>% 
  select(starts_with("auc"))

colnames(auc_mat_nhanes) <- c("8am", "4pm")
```

```{r auc_GLMMadaptive_nhanes, eval=FALSE}
auc_mat_nhanes2 <- nhanes_pred_adglmm %>%
  mutate(window = cut(t, breaks = window3, include.lowest = T)) %>% 
  select(Y, starts_with("pred"), window) %>%
  group_by(window) %>%
  summarise(auc1 = get_auc(Y, pred480),
            auc2 = get_auc(Y, pred960)) %>%
  filter(window != "[0,480]") %>% 
  select(starts_with("auc"))

colnames(auc_mat_nhanes2) <- c("8am", "4pm")
```

```{r, eval=FALSE}
data.frame(Window = c("8am-4pm", "4am-midnight"),
           auc_mat_nhanes, auc_mat_nhanes2, check.names = F) %>%
  kable(digits = 3, caption = "Area Under the ROC curve", booktabs = T) %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" "=1, "fGFPCA" = 2, "GLMMadaptive" = 2)) %>%
  add_header_above(c(" "= 1, "Maximum observation time" = 4))
```

@@

```{r pred_nhanes, fig.height=4, fig.width=10, eval=FALSE}
rand_id3 <- sample(unique(nhanes_pred_fgfpca$id), 4)

ggarrange(
  nhanes_pred_fgfpca %>% 
    filter(id %in% rand_id3) %>%
    mutate_at(vars(starts_with("pred")), function(x)exp(x)/(1+exp(x))) %>% 
    ggplot()+
    geom_line(aes(x=t, y = pred480, col = "8am"), linetype = "dashed", na.rm = T)+
    geom_line(aes(x=t, y = pred960, col = "4pm"),linetype = "dashed", na.rm = T)+
    geom_point(aes(x=t, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "fGFPCA")+
  scale_x_continuous(breaks = seq(0, 1440, by = 480), 
                     labels = c("Midnight", "8am", "4pm", "Midnight"))+
  theme(axis.text = element_text(size = 5)),

  nhanes_pred_adglmm %>% 
    filter(id %in% rand_id3) %>%
    mutate_at(vars(starts_with("pred")), function(x)exp(x)/(1+exp(x))) %>% 
    ggplot()+
    geom_line(aes(x=t, y = pred480, col = "8am"), linetype = "dashed", na.rm = T)+
    geom_line(aes(x=t, y = pred960, col = "4pm"),linetype = "dashed", na.rm = T)+
    geom_point(aes(x=t, y = Y, col = "Outcome"), size = 0.2)+
    facet_wrap(~id)+
    labs(x = "Time", y = "GLMMadaptive")+
  scale_x_continuous(breaks = seq(0, 1440, by = 480), 
                     labels = c("Midnight", "8am", "4pm", "Midnight"))+
  theme(axis.text = element_text(size = 5)),
  nrow = 1, common.legend = T)
```

>>>>

---
title: "Manuscript progress report"
author: "Ying Jin"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
    keep_tex: yes
  html_document:
    self_contained: no
    number_sections: yes
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float: yes
    font: 12pt
    keep_tex: yes
---

@@

bind_cols(fgfpca1, gfosr_l5_1, gfosr_l1_1, adglmm1, .name_repair = "minimal") %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" " = 1, "fGFPCA" = 4, "GFOSR (L=5)" = 4, 
                     "GFOSR (L=1)" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" " = 1, "Maximum observation time" = 16)) %>%
  group_rows(index = c("ISE"= 4, "AUC" = 4)) %>%
  landscape()

@@

bind_cols(fgfpca2, gfosr_l5_2, gfosr_l1_2, adglmm2, .name_repair = "minimal") %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" " = 1, "fGFPCA" = 4, "GFOSR (L=5)" = 4, 
                     "GFOSR (L=1)" = 4, "GLMMadaptive" = 4)) %>%
  add_header_above(c(" " = 1, "Maximum observation time" = 16)) %>%
  group_rows(index = c("ISE"= 4, "AUC" = 4)) %>%
  landscape()

@@

We take 60% subjects (N = 5257) for training, 40% (n = 3506) for out-of-sample prediction. When using the full sample The debiase step (mgcv::bam) in fGFPCA took a really a long time, and GLMM adaptive only works for intercept-only model. In the debias step, re-evaluated eigenvalues also lost their decreasing nature.

```{r load_nhanes}
# df_nhanes <- read_rds(here("Data/nhanes_bi.rds"))
load(here("Data/ApplOutput_fGFPCA.RData"))
load(here("Data/ApplOutput_GLMMadaptive.RData"))

load(here("Data/Appl_GFOSR_L5.RData"))
load(here("Data/Appl_GFOSR_L1.RData"))

pred_appl_gfosr_l1 <- pred_appl_gfosr_l1 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)
pred_appl_gfosr_l5 <- pred_appl_gfosr_l5 %>%
  rename(pred360=pred_w1, pred720=pred_w2, pred1080=pred_w3)

@@

```{r}
# window
window <- seq(0, 1440, by = 360)
```

```{r auc_appl}
auc_appl_fgfpca <- pred_nhanes_fgfpca %>% 
    mutate(window = cut(sind, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred360),
              auc2 = get_auc(Y, pred720),
              auc3 = get_auc(Y, pred1080)) %>%
    filter(window != "[0,360]") %>% 
    select(starts_with("auc"))

auc_appl_gfosr_l5 <- pred_appl_gfosr_l5 %>% 
  select(-window) %>%
    mutate(window = cut(sind, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred360),
              auc2 = get_auc(Y, pred720),
              auc3 = get_auc(Y, pred1080)) %>%
    filter(window != "[0,360]") %>% 
    select(starts_with("auc"))

auc_appl_gfosr_l1 <- pred_appl_gfosr_l1 %>% 
  select(-window) %>%
    mutate(window = cut(sind, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred360),
              auc2 = get_auc(Y, pred720),
              auc3 = get_auc(Y, pred1080)) %>%
    filter(window != "[0,360]") %>% 
    select(starts_with("auc"))

auc_appl_adglmm <- pred_nhanes_adglmm %>%
    mutate(window = cut(sind, breaks = window, include.lowest = T)) %>% 
    select(Y, starts_with("pred"), window) %>%
    group_by(window) %>%
    summarise(auc1 = get_auc(Y, pred360),
              auc2 = get_auc(Y, pred720),
              auc3 = get_auc(Y, pred1080)) %>%
    filter(window != "[0,360]") %>% 
    select(starts_with("auc"))

colnames(auc_appl_fgfpca) <- colnames(auc_appl_gfosr_l5) <- colnames(auc_appl_gfosr_l1) <- colnames(auc_appl_adglmm) <- c("6am", "12pm", "6pm")
```

```{r table_appl}
bind_cols(Window = c("6am-12pm", "12pm-6pm", "6pm-12am"),
          auc_appl_fgfpca, auc_appl_gfosr_l5, auc_appl_gfosr_l1, auc_appl_adglmm,
          .name_repair = "minimal") %>%
  kable(digits = 3, booktabs=T,
        table.attr="style=\"color:black;\"") %>%
  kable_styling(full_width = F) %>% 
  add_header_above(c(" " = 1, "fGFPCA" = 3, "GFOSR (L=5)" = 3, 
                     "GFOSR (L=1)" = 3, "GLMMadaptive" = 3)) %>%
  add_header_above(c(" " = 1, "Maximum observation time" = 12)) %>%
  landscape()
```

@@

```{r fig_appl, fig.height=8, fig.width=8}
rand_id <- sample(unique(pred_nhanes_adglmm$id), 4)
cols <- cols[1:3]
names(cols) <- c("6am", "12pm", "6pm")
bind_rows(
  pred_nhanes_fgfpca %>% filter(id %in% rand_id) %>% mutate(method="fGFPCA"),
  pred_appl_gfosr_l5 %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=5)"),
  pred_appl_gfosr_l1 %>% filter(id %in% rand_id) %>% mutate(method = "GFOSR (L=1)"),
  pred_nhanes_adglmm %>% filter(id %in% rand_id) %>% mutate(method = "GLMMadaptive")) %>%
  mutate_at(vars(pred360, pred720, pred1080), 
                .funs = function(x){exp(x)/(1+exp(x))}) %>%
  mutate(method=factor(method, 
         levels = c("fGFPCA", "GFOSR (L=5)", "GFOSR (L=1)","GLMMadaptive"))) %>%
  ggplot()+
  geom_point(aes(x=sind, y=Y), size = 0.2)+
  geom_line(aes(x=sind, y=pred360, col = "6am"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=sind, y=pred720, col = "12pm"), linetype="dashed", na.rm=T)+
  geom_line(aes(x=sind, y=pred1080, col = "6pm"), linetype="dashed", na.rm=T)+
  facet_grid(rows = vars(id), cols = vars(method))+
  labs(col = "Maximum observation time", x = "Time", y="")+
  scale_x_continuous(breaks = seq(0, 1440, by = 360),
                     labels = c("0am", "6am", "12pm", "6pm", "12am"))+
  theme(legend.position = "bottom", axis.text.x = element_text(size = 7))+
  scale_color_manual(values = cols)
```
